Hou10ni

### 2017-08-30 hou10ni/Tutorial/hou10ni_profile.tex
Hou10ni: kernel update

\lstset{language=FORTRAN}
\begin{lstlisting}[caption=kernel.f90]
SUBROUTINE fluid_inner_elt_ref(Nflu_inner, Nflusol_inner, nb_rhs, idx_vec_flu, idx_mat_flu, P_new, P_inter, P_old, A_flu, afluSize, pSize, vecSize_1, vecSize_2, matSize)
  implicit none
  INTEGER, PARAMETER :: dq=8 !< quadruple precision
  INTEGER, PARAMETER :: dp=4  !< double precision

@@

\begin{lstlisting}[caption=sub\_time\_loop.f90]
SUBROUTINE sub_time_loop
...
CALL fluid_inner_elt(Nflu_inner, Nflusol_inner, nb_rhs, idx_vec_flu, idx_mat_flu, P_new, P_inter, P_old, A_flu, afluSize, pSize, vecSize_1, vecSize_2, matSize)
...
END SUBROUTINE 
\end{lstlisting}

@@

We generate the exact same kernel as a reference version and run it with BOAST in order to compare the results in term of correctness and performances. KRef defines the reference version. This version simply needs to keep the Fortran code (referred as \texttt{codeF}) and wrap it with BOAST. To put it in a nutshell, a reference version should follow the template bellow.

\lstset{language=RUBY}
\begin{lstlisting}[caption=KRef\_template.rb]
def KRef
  lang = BOAST::get_lang
  BOAST::set_lang(BOAST::FORTRAN)
  kernel = BOAST::CKernel::new
  function_name = "function_ref"

  VARIABLES HERE
  FOR EXAMPLE : 
  geocode = BOAST::Int("geocode", :dir => :in )
  n01 = BOAST::Int("n01", :dir => :in)
  u = BOAST::Real("u", :dir => :in, :dim => [ BOAST::Dim(0, n01),  BOAST::Dim(0, n02))] )

   p = BOAST::Procedure::new(function_name, [geocode,n01,u])

   kernel.code.print <<EOF
  ORIGINAL C or FORTRAN CODE HERE
EOF

@@

 def initialize(options)
  @opts = {:preprocessor => false, :unroll => false, :inline => :included}
  @opts.update(options)
  # Parameters 
  @Nflu_inner    = Int("Nflu_inner", :dir => :in )
  @Nflusol_inner = Int("Nflusol_inner", :dir => :in )
  @nb_rhs        = Int("nb_rhs", :dir => :in )
  @afluSize      = Int("afluSize", :dir => :in )
  @pSize         = Int("pSize", :dir => :in )
  @vecSize_1     = Int("vecSize_1", :dir => :in )
  @vecSize_2     = Int("vecSize_2", :dir => :in )
  @matSize       = Int("matSize", :dir => :in )
  @idx_vec_flu   = Int("idx_vec_flu", :dir => :in, :dim => [Dim(@vecSize_1),Dim(@vecSize_2)] )
  @idx_mat_flu   = Int("idx_mat_flu", :dir => :in, :dim => [Dim(@matSize)] )
  @P_new         = Real("P_new",:dir => :inout, :dim => [Dim(@pSize),Dim(@nb_rhs)])
  @P_inter       = Real("P_inter",:dir => :in, :dim => [Dim(@pSize),Dim(@nb_rhs)])
  @P_old         = Real("P_old",:dir => :in, :dim => [Dim(@pSize),Dim(@nb_rhs)])
  @A_flu         = Real("A_flu",:dir => :in, :dim => [Dim(@afluSize)])
 end

@@

  push_env(:lang => FORTRAN) {
     @kernel = CKernel:: new()
     @kernel.procedure = Procedure("fluid_inner_elt_ref",[@Nflu_inner, @Nflusol_inner, @nb_rhs, @idx_vec_flu, @idx_mat_flu, @P_new, @P_inter, @P_old, @A_flu, @afluSize, @pSize,@vecSize_1,@vecSize_2,@matSize] ,:functions => nil)
     get_output.print codeF
  }
   return @kernel
 end

@@

# Creating ref kernel
k_ref_params = {:kernel => :ref, :LDFLAGS => "-lgfortran", :FCFLAGS => "-fbounds-check -O2"}
k = KRef::new(k_ref_params)
k.generate
puts k.kernel
k.kernel.build(:LDFLAGS => k_ref_params[:LDFLAGS], :FCFLAGS => k_ref_params[:FCFLAGS] )

# Parameters for loop iteration 1 and MPI Process 0 
# in the directory fluid_inner_elt_ref/sub_time_loop_1/
inputs = k.kernel.load_ref_inputs()
puts "** inputs =", inputs

inputs.each_key { |key|
  puts k.kernel.run(*(inputs[key]))
  puts k.kernel.compare_ref(outputs[key], inputs[key]).inspect
}

@@

\begin{lstlisting}[caption=Results (obtained with the command ruby run\_ref.rb)]
SUBROUTINE fluid_inner_elt_ref(Nflu_inner, Nflusol_inner, nb_rhs, idx_&
&vec_flu, idx_mat_flu, P_new, P_inter, P_old, A_flu, afluSize, pSize, v&
&ecSize_1, vecSize_2, matSize)
  implicit none
  INTEGER, PARAMETER :: dq=8 !< quadruple precision
  INTEGER, PARAMETER :: dp=4  !< double precision

@@

    @kernel = CKernel:: new()
    @kernel.procedure = Procedure("fluid_inner_elt_boast",[@Nflu_inner, @Nflusol_inner, @nb_rhs, @idx_vec_flu, @idx_mat_flu, @P_new, @P_inter, @P_old, @A_flu, @afluSize, @pSize,@vecSize_1,@vecSize_2,@matSize] ,:functions => nil)

@@

\begin{lstlisting}[caption=run\_boast.rb]
require 'BOAST'
require './KBoast.rb'
require 'narray_ffi'

set_default_real_size(4)

k_boast_params = {:kernel => :boast, :LDFLAGS => "-lgfortran", :FCFLAGS => "-fbounds-check -fimplicit-none -O2"}

set_lang(FORTRAN)
stats = []
repeat = 5

@@

inputs = k.kernel.load_ref_inputs()
outputs = k.kernel.load_ref_outputs()

inputs.each_key { |key|
  repeat.times {
    stats.push k.kernel.run(*(inputs[key]))
  }
  puts k.kernel.compare_ref(outputs[key], inputs[key]).inspect
}

@@

puts "#{k.kernel.procedure.name}: optim_nested = #{k_boast_params[:optim_nested]}  #{stats[:duration]} s ->  #{stats[:duration]*1.0e3} ms"
\end{lstlisting}

\begin{lstlisting}[caption=Results (obtained with the command ruby run\_boast.rb)]
SUBROUTINE fluid_inner_elt_boast(Nflu_inner, Nflusol_inner, nb_rhs, idx_&
&vec_flu, idx_mat_flu, P_new, P_inter, P_old, A_flu, afluSize, pSize, v&
&ecSize_1, vecSize_2, matSize)

  integer, parameter :: wp=kind(1.0d0)
  integer(kind=4), intent(in) :: Nflu_inner
  integer(kind=4), intent(in) :: Nflusol_inner
  integer(kind=4), intent(in) :: nb_rhs
  ...

END SUBROUTINE fluid_inner_elt_boast

{"P_new"=>0.0}
[{:duration=>0.011553553000000001, :start=>-8096482666880853504, :end=>-8084929113880853504}, {:duration=>0.011923121, :start=>-8122491033880853504, :end=>-8110567912880853504}, {:duration=>0.013262399000000001, :start=>-8135791916880853504, :end=>-8122529517880853504}, {:duration=>0.013991319, :start=>-8110525936880853504, :end=>-8096534617880853504}, {:duration=>0.014984313, :start=>-8150824959880853504, :end=>-8135840646880853504}]
fluid_inner_elt_boast: optim_nested = 3  0.011553553000000001 s ->  11.553553 ms
\end{lstlisting}

@@

Unroll both loops.

@@

    ## Procedure
    optim_2[@opts[:optim_main]].each { |f_main|
        pr f_main
    }

@@

\section{Comparison between MPI and MPI+OpenMP versions of hou10ni}
TODO

>>>>

\lstset{language=FORTRAN}
\begin{lstlisting}[caption=kernel.f90]
SUBROUTINE fluid_inner_elt_ref(Nflu_inner, Nflusol_inner, nb_rhs, afluSize, pSize, vecSize_1, vecSize_2, matSize, idx_vec_flu, idx_mat_flu, P_new, P_inter, P_old, A_flu)
  implicit none
  INTEGER, PARAMETER :: dq=8 !< quadruple precision
  INTEGER, PARAMETER :: dp=4  !< double precision

@@

\begin{lstlisting}[caption=sub\_time\_loop.f90]
SUBROUTINE sub_time_loop
...
CALL fluid_inner_elt(Nflu_inner, Nflusol_inner, nb_rhs, afluSize, pSize, vecSize_1, vecSize_2, matSize, idx_vec_flu, idx_mat_flu, P_new, P_inter, P_old, A_flu)
...
END SUBROUTINE 
\end{lstlisting}

@@

We generate the same kernel as a reference version (KRef ) and run it with BOAST in order to compare the results in term of correctness and performances. This version simply needs to keep the Fortran code (referred as \texttt{codeF}) and wrap it with BOAST. A reference version should follow the template bellow.

\lstset{language=RUBY}
\begin{lstlisting}[caption=KRef\_template.rb]
def KRef
  lang = BOAST::get_lang
  BOAST::set_lang(BOAST::FORTRAN)
  kernel = BOAST::CKernel::new
  kernel_name = "function_ref"

  # ARGUMENTS OF KERNEL HERE
  # FOR EXAMPLE : 
  geocode = BOAST::Int("geocode", :dir => :in )
  u = BOAST::Real("u", :dir => :in, :dim => [ BOAST::Dim(0, n01),  BOAST::Dim(0, n02))] )

   p = BOAST::Procedure::new(kernel_name, [geocode,u])

   kernel.code.print <<EOF
  # ORIGINAL C or FORTRAN CODE HERE
EOF

@@

 def initialize(options)
  # Parameters 
  @Nflu_inner    = Int("Nflu_inner", :dir => :in )
  @Nflusol_inner = Int("Nflusol_inner", :dir => :in )
  @nb_rhs        = Int("nb_rhs", :dir => :in )
  @afluSize      = Int("afluSize", :dir => :in )
  @pSize         = Int("pSize", :dir => :in )
  @vecSize_1     = Int("vecSize_1", :dir => :in )
  @vecSize_2     = Int("vecSize_2", :dir => :in )
  @matSize       = Int("matSize", :dir => :in )
  @idx_vec_flu   = Int("idx_vec_flu", :dir => :in, :dim => [Dim(@vecSize_1),Dim(@vecSize_2)] )
  @idx_mat_flu   = Int("idx_mat_flu", :dir => :in, :dim => [Dim(@matSize)] )
  @P_new         = Real("P_new",:dir => :inout, :dim => [Dim(@pSize),Dim(@nb_rhs)])
  @P_inter       = Real("P_inter",:dir => :in, :dim => [Dim(@pSize),Dim(@nb_rhs)])
  @P_old         = Real("P_old",:dir => :in, :dim => [Dim(@pSize),Dim(@nb_rhs)])
  @A_flu         = Real("A_flu",:dir => :in, :dim => [Dim(@afluSize)])
 end

@@

  push_env(:lang => FORTRAN) {
     @kernel = CKernel:: new()
     @kernel.procedure = Procedure("fluid_inner_elt_ref",[@Nflu_inner, @Nflusol_inner, @nb_rhs, @afluSize, @pSize,@vecSize_1,@vecSize_2,@matSize, @idx_vec_flu, @idx_mat_flu, @P_new, @P_inter, @P_old, @A_flu] ,:functions => nil)
     get_output.print codeF
  }
   return @kernel
 end

@@

# Creating ref kernel
k_ref_params = {:kernel => :ref, :LDFLAGS => "-lgfortran", :FCFLAGS => "-O2"}
k = KRef::new(k_ref_params)
k.generate
puts k.kernel
k.kernel.build(:LDFLAGS => k_ref_params[:LDFLAGS], :FCFLAGS => k_ref_params[:FCFLAGS] )

# Input parameters 
# found in the directory fluid_inner_elt_ref/sub_time_loop_1/ for example (created by the user)
inputs = k.kernel.load_ref_inputs()
puts "** inputs =", inputs

# Launch the kernel fluid_inner_elt_ref
# As many key as directories in 
inputs.each_key { |key|
  puts k.kernel.run(*(inputs[key]))
  puts k.kernel.compare_ref(outputs[key], inputs[key]).inspect
}

@@

\begin{lstlisting}[caption=Results (obtained with the command ruby run\_ref.rb)]
SUBROUTINE fluid_inner_elt_ref(Nflu_inner, Nflusol_inner, nb_rhs, afluSize, pSize, vecSize_1, &
&vecSize_2, matSize, idx_vec_flu, idx_mat_flu, P_new, P_inter, P_old, A_flu)
  implicit none
  INTEGER, PARAMETER :: dq=8 !< quadruple precision
  INTEGER, PARAMETER :: dp=4  !< double precision

@@

    @kernel = CKernel:: new()
    @kernel.procedure = Procedure("fluid_inner_elt_boast",[@Nflu_inner, @Nflusol_inner, @nb_rhs, @afluSize, @pSize,@vecSize_1,@vecSize_2,@matSize, @idx_vec_flu, @idx_mat_flu, @P_new, @P_inter, @P_old, @A_flu] ,:functions => nil)

@@

\begin{lstlisting}[caption=run\_boast.rb]
require 'BOAST'
require './KBoast.rb'
require 'narray_ffi'
include BOAST

k_boast_params = {:kernel => :boast, :LDFLAGS => "-lgfortran", :FCFLAGS => "-O2"}

set_default_real_size(4)
set_lang(FORTRAN)
stats = []
repeat = 5

@@

# Get input and output parameters # found in the directory fluid_inner_elt_boast/
inputs = k.kernel.load_ref_inputs()
outputs = k.kernel.load_ref_outputs()

# Launch the kernel
inputs.each_key { |key|
  repeat.times {
    stats.push k.kernel.run(*(inputs[key]))
  }
  puts k.kernel.compare_ref(outputs[key], inputs[key]).inspect
}

@@

# Output
puts "#{k.kernel.procedure.name}: optim_nested = #{k_boast_params[:optim_nested]}  #{stats[:duration]} s ->  #{stats[:duration]*1.0e3} ms"
\end{lstlisting}

\begin{lstlisting}[caption=Results (obtained with the command ruby run\_boast.rb)]
SUBROUTINE fluid_inner_elt_boast(Nflu_inner, Nflusol_inner, afluSize, pSize, vecSize_1, &
&vecSize_2, matSize, nb_rhs, idx_vec_flu, idx_mat_flu, P_new, P_inter, P_old, A_flu)
  integer, parameter :: wp=kind(1.0d0)
  integer(kind=4), intent(in) :: Nflu_inner
  integer(kind=4), intent(in) :: Nflusol_inner
  integer(kind=4), intent(in) :: nb_rhs
  ... 
END SUBROUTINE fluid_inner_elt_boast

{"P_new"=>0.0}  # No difference with the output

[{:duration=>0.011553553000000001, :start=>-8096482666880853504, :end=>-8084929113880853504}, {:duration=>0.011923121, :start=>-8122491033880853504, :end=>-8110567912880853504}, {:duration=>0.013262399000000001, :start=>-8135791916880853504, :end=>-8122529517880853504}, {:duration=>0.013991319, :start=>-8110525936880853504, :end=>-8096534617880853504}, {:duration=>0.014984313, :start=>-8150824959880853504, :end=>-8135840646880853504}]
fluid_inner_elt_boast: optim_nested = 3  0.011553553000000001 s ->  11.553553 ms
\end{lstlisting}

@@

The following code shows how we optimized the kernel with loop unrolling and using OpenMP.

@@

    ## Procedure
    optim_2[@opts[:optim_main]].each { |f_main|
        pr OpenMP::Parallel(:num_threads => @opts[:omp_num_threads]){
        pr OpenMP::For(:schedule => "static")
        pr f_main
        }
    }

@@

\subsubsection{Results}

The results were obtained using a Macbook Pro Intel Core i7 and the GNU fortran compiler. Values correspond to the minimum value of 20. Optimization flag: -O2
When \texttt{unroll main} and \texttt{unroll nested} are equal to 1, this means no unrolling is done.

\begin{table}[ht!]
\centering
\caption{Time depending on the number of OpenMP threads used for the main loop. No loop unrollment.}
{\scriptsize
\begin{tabular}{|c|c|c|} \hline 
\multirow{2}{*}{ {{\bf{Kernel vers.}}} }	 	  &  {{\bf{\# Threads}}}    &  \multirow{2}{*}{ {{\bf{Time (ms)}}} }  \\ 
					 		  & 	 	[1..6]   &  				\\ \hline \hline
ref  &         -  &  {\bf{83,769018}} \\ \hline
boast &     1 & 90,608141 \\ \hline
boast &      2 & 87,766917 \\  \hline
boast &    3 & 90,69217800 \\ \hline
boast &     4 & 87,29976 \\  \hline
boast &    5 & 84,527453 \\  \hline
boast &   6 & {\bf{81,92376500}} \\ \hline
\end{tabular}
}
\label{tab:1}
\end{table}

\begin{table}[ht!]
\centering
\caption{Statistics on time depending on the number of iterations unrolled for the main loop.}
{\scriptsize
\begin{tabular}{|c|c|c|c|c|} \hline 
\multirow{2}{*}{ {{\bf{Kernel vers.}}} }	 &   {{\bf{unroll main}}}   & {{\bf{unroll nested}}}   &    \multirow{2}{*}{ {{\bf{\# Threads}}} }  &  \multirow{2}{*}{ {{\bf{Time (ms)}}} }  \\ 
							 &    	[1..3]		 	    &   [1..10]	 		 & 							      &  				\\ \hline \hline
boast &    1 &  &  & 90,608141 \\ \cline{1-2}  \cline{5-5}
boast &    2 & 1 & 1 & 88,888357 \\  \cline{1-2}  \cline{5-5}
boast &    3 &  & & 89,98507500 \\ \hline \hline

boast &   1 &  &  & 88,516961 \\ \cline{1-2}  \cline{5-5}
boast &    2 & 2 & 1 & 87,892416 \\ \cline{1-2}  \cline{5-5}
boast &    3 &  &  & 86,117465 \\ \hline \hline

boast &    1 &  &  & 86,232686 \\ \cline{1-2}  \cline{5-5}
boast &    2 & 3 & 1 & 86,40083400 \\ \cline{1-2}  \cline{5-5}
boast &   3 &  &  & 82,334101 \\ \hline \hline

boast &    1 &  &  & 88,345071000 \\ \cline{1-2}  \cline{5-5}
boast &    2 & 4 & 1 & 82,89916300 \\ \cline{1-2}  \cline{5-5}
boast &    3 &  &  & 82,646548 \\ \hline \hline

boast &    1 & &  & 91,76817400 \\ \cline{1-2}  \cline{5-5}
boast &    2 & 5 & 1 & 89,60873700 \\ \cline{1-2}  \cline{5-5}
boast &   3 &  &  & 83,658755 \\ \hline \hline

boast &   1 & &  & 92,428135000 \\ \cline{1-2}  \cline{5-5}
boast &    2 & 6 & 1 & 84,5453400 \\ \cline{1-2}  \cline{5-5}
boast &  3 &  &  & 87,82848900 \\ \hline \hline

boast &    1 &  &  & 85,81414200 \\ \cline{1-2}  \cline{5-5}
boast &    2 & 7 & 1 & 89,83438 \\ \cline{1-2}  \cline{5-5}
boast &   3 &  & & {\bf{81,77698900}}  \\ \hline \hline

boast &    1 &  &  & 85,221406 \\ \cline{1-2}  \cline{5-5}
boast &    2 & 8 & 1 & 85,728041 \\ \cline{1-2}  \cline{5-5}
boast &   3 &  &  & 87,331661 \\ \hline \hline

boast &   1 &  &  & 88,48286500 \\ \cline{1-2}  \cline{5-5}
boast &    2 & 9 & 1 & 87,05563300 \\ \cline{1-2}  \cline{5-5}
boast &    3 &  &  & 86,669093 \\ \hline \hline

boast &    1 &  &  & 92,284591 \\ \cline{1-2}  \cline{5-5}
boast &    2 & 10 & 1 & 85,477758 \\ \cline{1-2}  \cline{5-5}
boast &   3 & &  & 85,488478 \\ \hline
\end{tabular}
}
\label{tab:2}
\end{table}

\begin{table}[ht!]
\centering
\caption{Time depending on the number of iterations unrolled for the nested loop.}
{\scriptsize
\begin{tabular}{|c|c|c|c|c|} \hline 
\multirow{2}{*}{ {{\bf{Kernel vers.}}} }	 	 &  \multirow{2}{*}{ {{\bf{unroll main}}} }  &  {{\bf{unroll nested}}}   &  \multirow{2}{*}{ {{\bf{\# Threads}}} }   &  \multirow{2}{*}{ {{\bf{Time (ms)}}} }  \\ 
						 		 &   		 					    &  	 		[1..10]				           & 	 						      &  				\\ \hline \hline
boast &     & 1 &  & 90,608141 \\ \cline{1-1}  \cline{3-3}  \cline{5-5}
boast &     & 2 &  & 88,516961 \\ \cline{1-1}  \cline{3-3}  \cline{5-5}
boast &    & 3 &  & 86,232686 \\ \cline{1-1}  \cline{3-3}  \cline{5-5} 
boast &   & 4 &  & 88,345071000 \\ \cline{1-1}  \cline{3-3}  \cline{5-5}
boast &   1 & 5 & 1 & 91,76817400 \\ \cline{1-1}  \cline{3-3}  \cline{5-5}
boast &     & 6 &  & 92,42813500 \\ \cline{1-1}  \cline{3-3}  \cline{5-5}
boast &    & 7 &  & 85,81414200 \\ \cline{1-1}  \cline{3-3}  \cline{5-5}
boast &     & 8 &  & {\bf{85,221406}} \\ \cline{1-1}  \cline{3-3}  \cline{5-5}
boast &    & 9 &  & 88,48286500 \\ \cline{1-1}  \cline{3-3}  \cline{5-5}
boast &     & 10 &  & 92,284591 \\ \hline
\end{tabular}
}
\label{tab:3}
\end{table}

\begin{table}[ht!]
\centering
\caption{All values under the reference version.}
{\scriptsize
\begin{tabular}{|c|c|c|c|c|} \hline 
\multirow{2}{*}{ {{\bf{Kernel vers.}}} }	 	 &  \multirow{2}{*}{ {{\bf{unroll main}}} }  &  \multirow{2}{*}{ {{\bf{unroll nested}}} }  &  \multirow{2}{*}{ {{\bf{\# Threads}}} }   &  \multirow{2}{*}{ {{\bf{Time}}} }  \\ 
						 		 &    							    &  	 							  & 	 						      &  				\\ \hline \hline
ref    &   &   &   &  0.083769018 \\ \hline \hline

boast &    3 & 7 & 1 & {\bf{81,77698900}} \\ \hline
boast &   2 & 3 & 5 & 82,006891 \\ \hline
boast &   3 & 4 & 3 & 82,17091900\\ \hline
boast &   3 & 3 & 1 & 82,334101 \\ \hline
boast &   3 & 4 & 1 & 82,646548 \\ \hline
boast &    2 & 4 & 5 & 82,809038 \\ \hline
boast &    2 & 4 & 4 & 82,81120700 \\ \hline
boast &    3 & 3 & 2 & 82,818522 \\ \hline
boast  &    1 & 5 & 4 & 82,866756 \\ \hline
boast &    2 & 4 & 1 & 82,89916300 \\ \hline
boast &    3 & 2 & 3 & 82,949804 \\ \hline
boast  &    1 & 3 & 4 & 82,994519 \\ \hline
boast &    1 & 2 & 3 & 83,047285 \\ \hline
boast &    3 & 7 & 2 & 83,08922 \\ \hline
boast  &    2 & 3 & 6 & 83,09479700 \\ \hline
boast &    3 & 4 & 6 & 83,137816 \\ \hline
boast &    3 & 1 & 3 & 83,15446500 \\ \hline
boast &    3 & 5 & 4 & 83,33983 \\ \hline
boast &    2 & 7 & 6 & 83,558517 \\ \hline
boast &    2 & 7 & 2 & 83,57951500 \\ \hline
\end{tabular}
}
\label{tab:4}
\end{table}

%\section{Comparison between MPI and MPI+OpenMP versions of hou10ni}
%TODO
